---
title: "Normalization"
author: "Maggie Sha"
date: "3/18/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(plyr)
```

```{r echo=FALSE}
adult <- read.csv("EarthAgency_Adults_R.csv", header = TRUE)

children <- read.csv("EarthAgency_Children_R.csv", header = TRUE)


adult_df <- adult %>%

  dplyr::select(

    Condition,

    Agency_Language,

    SRFactsTotal,

    invitalscore,

    inpsychscore,

    MeanSever,

    BioJtscore,

    AntJtscore,

    BioJFtotal,

    FirstLang,

    Age

  ) %>%

  mutate(

    Condition = as.factor(Condition),

    SRFactsTotal = as.integer(SRFactsTotal),

    invitalscore = as.integer(invitalscore),

    inpsychscore = as.integer(inpsychscore),

    Agency_Language = as.factor(Agency_Language),

    FirstLang = as.factor(FirstLang),

    Age

  )







#compress scores to match invitalscore and inpsychscore for children
# 
# adult_df$invitalscore[which(adult_df$invitalscore==0 | adult_df$invitalscore==1)]<-0
# 
# adult_df$invitalscore[which(adult_df$invitalscore==2) ]<-1
# 
# adult_df$invitalscore[which(adult_df$invitalscore==3) ]<-2
# 
# adult_df$invitalscore[which(adult_df$invitalscore==4 | adult_df$invitalscore==5)]<-3
# 
# 
# 
# adult_df$inpsychscore[which(adult_df$inpsychscore==2 | adult_df$inpsychscore==3)]<-2
# 
# adult_df$inpsychscore[which(adult_df$inpsychscore==4)]<-3
# 
# adult_df$inpsychscore[which(adult_df$inpsychscore==5)]<-4



child_df <- children %>%

  dplyr::select(

    Condition,

    Agency_Language,

    SRFactsTotal,

    invitalscore,

    inpsychscore,

    MeanSever,

    BioJtscore,

    AntJtscore,

    BioJFtotal,

    Age

  ) %>%

  mutate(

    Condition = as.factor(Condition),

    SRFactsTotal = as.integer(SRFactsTotal),

    invitalscore = as.integer(invitalscore),

    inpsychscore = as.integer(inpsychscore),

    Agency_Language = as.factor(Agency_Language),

    Age 

  )



# 1 = speaking English as their first language

FirstLang <- rep(1, nrow(child_df))

child_df <- child_df %>% 

  mutate(

    FirstLang = as.factor(FirstLang),

    )




# invitalscore normalization
adult_min <- min(adult_df$inpsychscore)
adult_max <- max(adult_df$inpsychscore)
child_min <- min(child_df$inpsychscore)
child_max <- max(child_df$inpsychscore)
adult_df$normalized_inpsychscore <- (adult_df$inpsychscore - adult_min)/(adult_max - adult_min)
child_df$normalized_inpsychscore <- (child_df$inpsychscore - child_min)/(child_max - child_min)


# invitalscore normalization
adult_min2 <- min(adult_df$invitalscore)
adult_max2 <- max(adult_df$invitalscore)
child_min2 <- min(child_df$invitalscore)
child_max2 <- max(child_df$invitalscore)
adult_df$normalized_invitalscore <- (adult_df$invitalscore - adult_min2)/(adult_max2 - adult_min2)
child_df$normalized_invitalscore <- (child_df$invitalscore - child_min2)/(child_max2 - child_min2)


#To take care of the one errant entry, children[87] (mean severity scores are divided by 4 so .67 is not possible…)

child_df$MeanSever[child_df$MeanSever==2.67] <- 2.5



#To remove children[41] which has no BIoJtscore or AntJtscore

#To remove children[45] which has no Agency_Language or SRFactsTotal

child_df <- na.omit(child_df)



## COMBINING children and adult into one data.frame

AC_df <- rbind(adult_df, child_df)



#Giving factor levels cleaner language

AC_df$Condition <- revalue(AC_df$Condition, c("1"="Obj", "2"="Nat", "3"="Per"))

AC_df$Agency_Language <- revalue(AC_df$Agency_Language, c("0"="Obj", "2"="Nat", "3"="Per"))

AC_df$Age <- ifelse(AC_df$Age > 10, "Adult", "Child")

AC_df$Age <- factor(AC_df$Age)
```



# 1. Mean Severity Ordinal model with combined levels (bcm)
```{r message=FALSE}
#Meansever combined into three levels (bcm)
AC_df$comMeanSever <- AC_df$MeanSever
AC_df$comMeanSever[which((AC_df$MeanSever < 1.5))]<-1
AC_df$comMeanSever[which((AC_df$MeanSever > 1.25 & AC_df$MeanSever < 2.25))]<-2
AC_df$comMeanSever[which((AC_df$MeanSever > 2))]<-3
AC_df$comMeanSever <- factor(AC_df$comMeanSever, ordered = TRUE)

table(AC_df$MeanSever)
table(AC_df$comMeanSever)

plor.fit1c <- polr(comMeanSever ~  Condition + Agency_Language + SRFactsTotal + 
                     normalized_invitalscore  + normalized_inpsychscore + FirstLang+Age, data = AC_df)
#summary(plor.fit1c)

#Get p-value
ctable <- round(coef(summary(plor.fit1c)), 3)
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
ctable1 <- cbind(ctable, "p value" =round(p, 3))
ctable1

#Compute confusion table and misclassification error
fitted <- predict(plor.fit1c)
n <- table(AC_df$comMeanSever, fitted)
prop.correct <- c(n[1,1]/sum(n[1,]),n[2,2]/sum(n[2,]),n[3,3]/sum(n[3,]))
prop.correct <- round(100*prop.correct, 0)
rbind(n,prop.correct)
ME <- mean(as.character(AC_df$comMeanSever) !=as.character(fitted))
cat("Misclassification error is:",ME,"\n")
```


# 2. Biocentric Justification score, Ordinal model with combined levels (bcm)
```{r message=FALSE}
#BioJtscore with Combination (bcm)
AC_df$comBioJtscore<-AC_df$BioJtscore
AC_df$comBioJtscore[which((AC_df$comBioJtscore==0))]<-1
AC_df$comBioJtscore[which((AC_df$comBioJtscore==2))]<-1
AC_df$comBioJtscore[which((AC_df$comBioJtscore==3))]<-2
AC_df$comBioJtscore[which((AC_df$comBioJtscore==4))]<-3
AC_df$comBioJtscore<- factor(AC_df$comBioJtscore, ordered = TRUE)

table(AC_df$BioJtscore)
table(AC_df$comBioJtscore)

plor.fit2c <- polr(comBioJtscore ~  Condition + Agency_Language + 
                     SRFactsTotal + normalized_invitalscore  + normalized_inpsychscore + FirstLang + Age, data = AC_df)
#summary(plor.fit2c)

#Get p-value
ctable <- round(coef(summary(plor.fit2c)), 3)
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
ctable1 <- cbind(ctable, "p value" = round(p, 3))
ctable1

#Compute confusion table and misclassification error
fitted <- predict(plor.fit2c)
n <- table(AC_df$comBioJtscore, fitted)
prop.correct <- c(n[1,1]/sum(n[1,]),n[2,2]/sum(n[2,]),n[3,3]/sum(n[3,]))
prop.correct <- round(100*prop.correct, 0)
rbind(n,prop.correct)
ME <- mean(as.character(AC_df$comBioJtscore) !=as.character(fitted))
cat("Misclassification error is:",ME,"\n")
```



# 3. Anthropocentric Justification score, Ordinal model with combined levels (bcm)
```{r message=FALSE}
#AntJtscore with Combination (bcm)
AC_df$comAntJtscore<-AC_df$AntJtscore
AC_df$comAntJtscore[which((AC_df$comAntJtscore==0))]<-1
AC_df$comAntJtscore[which((AC_df$comAntJtscore==2))]<-1
AC_df$comAntJtscore[which((AC_df$comAntJtscore==3))]<-2
AC_df$comAntJtscore[which((AC_df$comAntJtscore==4))]<-3
AC_df$comAntJtscore<- factor(AC_df$comAntJtscore, ordered = TRUE)

table(AC_df$AntJtscore)
table(AC_df$comAntJtscore)

plor.fit3c<- polr(comAntJtscore~  Condition + Agency_Language + SRFactsTotal + 
                    normalized_invitalscore  + normalized_inpsychscore + FirstLang + Age, data = AC_df)
#summary(plor.fit3c)

#Get p-value
ctable <- round(coef(summary(plor.fit3c)), 3)
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
ctable1 <- cbind(ctable, "p value" = round(p, 3))
ctable1

#Compute confusion table and misclassification error
fitted <- predict(plor.fit3c)
n <- table(AC_df$comAntJtscore, fitted)
prop.correct <- c(n[1,1]/sum(n[1,]),n[2,2]/sum(n[2,]),n[3,3]/sum(n[3,]))
prop.correct <- round(100*prop.correct, 0)
rbind(n,prop.correct)
ME <- mean(as.character(AC_df$comAntJtscore) !=as.character(fitted))
cat("Misclassification error is:",ME,"\n")
```


# 4. Biocentric choice score, Ordinal model with combined levels (bcm)
```{r message=FALSE}
#BioJFtotal with Combination
AC_df$comBioJFtotal<-AC_df$BioJFtotal
AC_df$comBioJFtotal[which((AC_df$comBioJFtotal==0))]<-1
#AC_df$comBioJFtotal[which((AC_df$comBioJFtotal==2))]<-1
#AC_df$comBioJFtotal[which((AC_df$comBioJFtotal==3))]<-2
AC_df$comBioJFtotal[which((AC_df$comBioJFtotal==4))]<-3
AC_df$comBioJFtotal<- factor(AC_df$comBioJFtotal, ordered = TRUE)

table(AC_df$BioJFtotal)
table(AC_df$comBioJFtotal)

plor.fit4c <- polr(comBioJFtotal ~  Condition + Agency_Language + SRFactsTotal + 
                     normalized_invitalscore  + normalized_inpsychscore + FirstLang + Age, data = AC_df)
#summary(plor.fit4c)

#Get p-value
ctable <- round(coef(summary(plor.fit4c)), 3)
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
ctable1 <- cbind(ctable, "p value" = round(p, 3))
ctable1

#Compute confusion table and misclassification error
fitted <- predict(plor.fit4c)
n <- table(AC_df$comBioJFtotal, fitted)
prop.correct <- c(n[1,1]/sum(n[1,]),n[2,2]/sum(n[2,]),n[3,3]/sum(n[3,]))
prop.correct <- round(100*prop.correct, 0)
rbind(n,prop.correct)
ME <- mean(as.character(AC_df$comBioJFtotal) !=as.character(fitted))
cat("Misclassification error is:",ME,"\n")
```

# 5. Anova test comparing Age
```{r}
#Mean Severity
plor.fit1c2 <- polr(comMeanSever ~  Condition + Agency_Language + 
                      SRFactsTotal + normalized_invitalscore  + normalized_inpsychscore + FirstLang, data = AC_df)

anova(plor.fit1c, plor.fit1c2)


#Biocentric Justification score
plor.fit2c2 <- polr(comBioJtscore ~  Condition + Agency_Language + 
                      SRFactsTotal + normalized_invitalscore  + normalized_inpsychscore + FirstLang, data = AC_df)

anova(plor.fit2c, plor.fit2c2)


#Anthropocentric Justification score
plor.fit3c2<- polr(comAntJtscore ~ Condition + Agency_Language + 
                     SRFactsTotal + normalized_invitalscore  + normalized_inpsychscore + FirstLang, data = AC_df)

anova(plor.fit3c,plor.fit3c2)


#Biocentric choice score
plor.fit4c2 <- polr(comBioJFtotal ~  Condition + Agency_Language + SRFactsTotal + 
                      normalized_invitalscore  + normalized_inpsychscore + FirstLang, data = AC_df)

anova(plor.fit4c,plor.fit4c2)
```


# 6. Anova test comparing Condition
```{r}
#Mean Severity
plor.fit1c2 <- polr(comMeanSever ~ Age + Agency_Language + 
                      SRFactsTotal + normalized_invitalscore  + normalized_inpsychscore + FirstLang, data = AC_df)

anova(plor.fit1c, plor.fit1c2)


#Biocentric Justification score
plor.fit2c2 <- polr(comBioJtscore ~  Age + Agency_Language + 
                      SRFactsTotal + normalized_invitalscore  + normalized_inpsychscore + FirstLang, data = AC_df)

anova(plor.fit2c, plor.fit2c2)


#Anthropocentric Justification score
plor.fit3c2<- polr(comAntJtscore ~ Age + Agency_Language + 
                     SRFactsTotal + normalized_invitalscore  + normalized_inpsychscore + FirstLang, data = AC_df)

anova(plor.fit3c,plor.fit3c2)


#Biocentric choice score
plor.fit4c2 <- polr(comBioJFtotal ~ Age + Agency_Language + SRFactsTotal + 
                      normalized_invitalscore  + normalized_inpsychscore + FirstLang, data = AC_df)

anova(plor.fit4c,plor.fit4c2)
```


# 7. Anova test comparing normalization
```{r}
#Mean Severity
plor.fit1c2 <- polr(comMeanSever ~ Condition + Age + Agency_Language + 
                      SRFactsTotal + invitalscore  + inpsychscore + FirstLang, data = AC_df)

anova(plor.fit1c, plor.fit1c2)


#Biocentric Justification score
plor.fit2c2 <- polr(comBioJtscore ~  Condition + Age + Agency_Language + 
                      SRFactsTotal + invitalscore  + normalized_inpsychscore + FirstLang, data = AC_df)

anova(plor.fit2c, plor.fit2c2)


#Anthropocentric Justification score
plor.fit3c2<- polr(comAntJtscore ~ Condition + Age + Agency_Language + 
                     SRFactsTotal + invitalscore  + inpsychscore + FirstLang, data = AC_df)

anova(plor.fit3c,plor.fit3c2)


#Biocentric choice score
plor.fit4c2 <- polr(comBioJFtotal ~ Condition + Age + Agency_Language + SRFactsTotal + 
                      invitalscore  + inpsychscore + FirstLang, data = AC_df)

anova(plor.fit4c,plor.fit4c2)
```